---
title: "Outlier exploration"
author: "Álvaro Gómez Pérez"
date: "2025-01-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Packages
```{r}
library(dplyr)
library(data.table)
library(limma)
library(ggplot2)
library(patchwork)
library(tidyr)
library(tibble)
library(stringr)
library(gprofiler2)
library(xlsx)
library(stringr)
library(gridExtra)
library(grid)
suppressPackageStartupMessages(library(ComplexHeatmap))
suppressPackageStartupMessages(library(circlize))
library(factoextra)
library(glue)
library(dendextend)
library(ggrepel)
library(ggvenn)
library(tidyr)
library(diann)
library(platetools)
library(viridis)
```

Set parameters
```{r}
# Significance level to be used for all tests and plots in this file
alpha <- 0.01

# Log-fold change limit to be considered "biologically significant"
lfc_threshold <- 1.5

# Number of nDEP from which a tRNA is considered "major" (this threshold value should be included)
threshold_major <- 2

# Set directories to be used
working_from = "charite"

if (working_from == "home") {
  base_dir = "/home/alvaro/MyStuff/tRNA_KOs/"
} else
if (working_from == "charite") {
  base_dir = "C:/MyStuff/tRNA_KOs/"
}
```


# 1. Load data and initial analysis
## 1.1. Load data
```{r}
# Proteomics data
proteomics_raw <- read.delim2(paste(base_dir, 'tRNA_KOs/Data/proteomics_data/AlternativeAAUsage-tRNA_peptidecentric_ProteinMaxLFQ_driftcorrected_batchcorrected_onWTs.tsv', sep=""), 
                              header = T)

# Sample layout
sample_layout <- as.data.frame(fread(paste(base_dir, "tRNA_KOs/Data/basic/sample_layout_alvaro.tsv", sep="")))

# The master tRNA dataset
master_dataset <- as.data.frame(fread(paste(base_dir, "tRNA_KOs/Data/basic/master_tRNA_dataset.csv", sep="")))

# The ODs
od <- as.data.frame(fread(paste(base_dir, "tRNA_KOs/Data/basic/OD_final.csv", sep = "")))

# Phenotypic data
phenotypic_data <- as.data.frame(fread(paste(base_dir, "tRNA_KOs/Data/Articles/bloom_ackermann_2014/phenotypic_results_full.tsv", sep="")))

# Import the functions I have created in another file
source(paste(base_dir, "tRNA_KOs/Code/R/Mine/0.general_use_functions.R", sep=""))


# limpa output
dpcfit <- readRDS(paste(base_dir, "tRNA_KOs/Data/Other/limpa_analysis/dpcfit.rds", sep=""))
y.protein <- readRDS(paste(base_dir, "tRNA_KOs/Data/Other/limpa_analysis/y.protein.rds", sep=""))

## Set this so I don't have to change its name in the code below
trna_ko <- y.protein$E
```

## 1.2. Obtain processed protein dataset from the precursor-level processed data, so as to still have NAs
```{r}
# DONE ONCE, NOT NECESSARY TO RUN AGAIN SINCE IT TAKES A LONG TIME, JUST LOAD THE RESULTS

# Get report into long format
#report <- report %>%
#  pivot_longer(cols = !c(Precursor.Id, Stripped.Sequence, Protein.Group, Genes),
#               names_to = "File.Name",
#               values_to = "Precursor.Quantity")
#
## Turn negative values to NAs
#report <- report %>%
#  mutate(Precursor.Quantity = case_when(Precursor.Quantity < 0 ~ NA,
#                                        TRUE ~ Precursor.Quantity))

# Run MaxLFQ on this data
#proteomics_processed <- diann_maxlfq(report,
#                                     sample.header = "File.Name",                        # Column with the sample IDs
#                                     group.header = "Genes",                             # Column with the protein IDs
#                                     id.header = "Precursor.Id",                         # Column with the precursor IDs
#                                     quantity.header = "Precursor.Quantity")             # Column with the precursor quantities
#
#proteomics_processed <- data.frame(proteomics_processed)
#fwrite(proteomics_processed, paste(base_dir, "Data/Other/proteomics_data/protein_level_processed_not_imputed.tsv", sep=""))

# Load the resulting file from MaxLFQ
proteomics_processed <- as.data.frame(fread(paste(base_dir, "tRNA_KOs/Data/proteomics_data/protein_level_processed_not_imputed.tsv", sep="")))
```

## 1.3. Merge sample_layout with master_dataframe
```{r}
# Merge with master dataframe
colnames(master_dataset)[colnames(master_dataset) == "gene_name"] <- "Strain.Name"

sample_layout <- sample_layout %>% 
  dplyr::select(-Anticodon) %>%
  left_join(master_dataset, by = c("Strain.Name"))
```

## 1.4. Pre-process proteomics data - this is where I:
  - Deal with the 2/3 duplicated strains
  - Remove samples in the wrong subbatch (when dataset == "outlier_removed")
  - Generally do everything necessary to come up with trna_ko from proteomics_raw
```{r}
# Remove annoying unnecessary (?) column (UniProt IDs)
if (dataset == "outlier_removed") 
  {proteomics_raw <- proteomics_raw %>%
    dplyr::select(-sample) %>%
    rename(genes = X) %>%
    filter(!(genes %in% c('', 'Genes', 'Protein.Group'))) %>%                         # Remove rows for which "genes" column is empty or has "Genes" as value
    column_to_rownames(var = 'genes')}                                                # Convert gene name column to rownames

if (dataset == "batchcorrected_onWTs")    
    {proteomics_raw <- proteomics_raw %>%
    dplyr::select(-Protein.Group) %>%
      rename(genes = Genes) %>%
      column_to_rownames(var = 'genes')}


# Come up with new colnames
new_names <- c()
current_names <- colnames(proteomics_raw)
new_names <- c(new_names, as.character(sample_layout$final_proteomics_colnames[match(current_names, sample_layout$raw_proteomics_colnames)]))
colnames(proteomics_raw) <- new_names


# Fix the situation with 2 tRNA_KOs being present twice (so 6 replicates for each instead of 3) - just going to take the average between the replicated samples
## Come up with the averages
n_occur <- data.frame(table(colnames(proteomics_raw)))
repeated_samples <- as.character(n_occur$Var1)[n_occur$Freq > 1]

out_temp <- data.frame(matrix(ncol = 0, nrow = nrow(proteomics_raw)))
for (i in 1:length(repeated_samples)) {
  strain <- repeated_samples[i]
  temp <- proteomics_raw[, colnames(proteomics_raw) == strain]
  temp[,1] <- as.numeric(temp[,1])
  temp[,2] <- as.numeric(temp[,2])
  out <- c(apply(temp, 1, mean))
  out_temp[,i] <- out
}
colnames(out_temp) <- repeated_samples

## Remove original columns from dataframe and add the averaged ones
proteomics_raw <- proteomics_raw[,!(colnames(proteomics_raw) %in% repeated_samples)]
proteomics_raw <- cbind(proteomics_raw, out_temp)


# Format data to get a biological protein abundance dataframe with samples names as colnames and gene names as rownames
trna_ko <- proteomics_raw %>%                 
  dplyr::select(!contains('QC'))                                 # Remove QCs columns
                                

# Convert to a numeric dataframe
trna_ko <- as.data.frame(apply(trna_ko, 2, as.numeric), row.names =  rownames(trna_ko))   

# Remove the sub-batch that went wrong
if (dataset == "outlier_removed") {
  samples_in_wrong_batch <- sample_layout$final_proteomics_colnames[sample_layout$Analysis.Plate.96_Replicate == "3_2"]
trna_ko <- trna_ko[,!(colnames(trna_ko) %in% samples_in_wrong_batch)]
}

# ADD A STEP WHERE I TURN THE ROWNAMES TO ALL BE SYSTEMATIC GENE NAMES RATHER THAN STANDARD AND SYSTEMATIC MIXED?


# Save this last dataset for later usage elsewhere
fwrite(trna_ko, paste(base_dir, "tRNA_KOs/Data/proteomics_data/processed_proteomics_dataframe_batchcorrected_onWTs.tsv", sep=""), row.names = T)


# Remove unnecessary objects
rm(proteomics_raw, n_occur, repeated_samples, out_temp, strain, temp, out, i, new_names, samples_in_wrong_batch, current_names)
```

## 1.5. Come up with plate drawings for where the samples were located
```{r}
# 384-well plates
replicates <- c(1, 2, 3)
for (i in replicates) {
  temp <- sample_layout %>%
    filter(Replicate == i)
  
  raw_map(data = temp$Analysis.Plate.96,
        well = temp$Position.Within.Plate.384,
        plate = 384) +
  ggtitle(glue("384-well plate {i} - tRNA KOs")) +
  theme_dark() +
  scale_fill_viridis_d() +
  theme(axis.text=element_text(size=6))
  
  ggsave(paste(base_dir, "tRNA_KOs/Output/Plots/sample_layout/tRNA_KOs_384_well_plate_", i, ".png", sep=""))
}

# 96-well plates
for (i in replicates) {
  temp <- sample_layout %>%
    filter(Analysis.Plate.96 == i & Analysis.Plate.384 == i)        # I need to filter on 384-well plate as well so as not to have 3 plates here
  
  raw_map(data = temp$Strain.Type,
        well = temp$Position.Within.Plate.96,
        plate = 96) +
  ggtitle(glue("96-well plate {i} - tRNA KOs")) +
  theme_dark() +
  scale_fill_viridis_d() +
  theme(axis.text=element_text(size=6))
  
  ggsave(paste(base_dir, "tRNA_KOs/Output/Plots/sample_layout/tRNA_KOs_96_well_plate_", i, ".png", sep=""))
}
```




# 2. Check relationship between date and plate - it was plate 1 that was ran between 2 days, so that is not the cause of the weird outliers in plate 2 :(
```{r}
ggplot(data = sample_layout, aes(x = Replicate, y = date)) +
  geom_point()
```




# 3. PCA
First look at the PCA plot colored by all variables possible (both metadata and biological), see if there is any variable that explains the weird outliers in Batch 2. 

Prepare PCA data
```{r}
# Get data PCA-ready
PCA_data <- as.data.frame(t(trna_ko))

# Get PCA tags for different cases - define a function for this, then call it before running the corresponding plot
get_PCA_tags <- function(PCA_data, sample_layout, type_of_tags) {
  out <- c()
  for (i in 1:nrow(PCA_data)) {
    rowname <- rownames(PCA_data)[i]
    out <- c(out, sample_layout[sample_layout$final_proteomics_colnames == rowname, type_of_tags][1])
  }
  return(out)
}

# FOR THESE ONES TO WORK AGAIN I WILL PROBABLY HAVE TO MERGE SAMPLE_LAYOUT AND MASTER_DATASET, PROBABLY EASY
PCA_tags_aa <- get_PCA_tags(PCA_data, sample_layout, "AA")
PCA_tags_chr <- get_PCA_tags(PCA_data, sample_layout, "chromosome")
PCA_tags_anticodon <- get_PCA_tags(PCA_data, sample_layout, "anticodon")
```

Run PCA
```{r}
#calculate principal components
results_PCA <- prcomp(PCA_data, scale = TRUE)

#reverse the signs
results_PCA$rotation <- -1*results_PCA$rotation

#calculate total variance explained by each principal component
var_explained =  results_PCA$sdev^2 / sum(results_PCA$sdev^2)
var_exp_PC1 = format(round(var_explained[1]*100, 3), nsmall = 3)
var_exp_PC2 = format(round(var_explained[2]*100, 3), nsmall = 3)
```

Plot PCA - by metadata
```{r}
# Label by batch
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "Replicate")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by batch") +
  theme_light()

# Label by date
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "date")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by date") +
  theme_light()


# Label by Analysis.Plate.384
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "Analysis.Plate.384")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by Analysis.Plate.384") +
  theme_light()


# Label by Analysis.Column.384
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "Analysis.Column.384")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by Analysis.Column.384") +
  theme_light()


# Label by Analysis.Row.384
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "Analysis.Row.384")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by Analysis.Row.384") +
  theme_light()


# Label by Analysis.Plate.96
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "Analysis.Plate.96")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by Analysis.Plate.96") +
  theme_light()


# Label by Analysis.Row.96
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "Analysis.Row.96")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by Analysis.Row.96") +
  theme_light()


# Label by Analysis.Column.96
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "Analysis.Column.96")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by Analysis.Column.96") +
  theme_light()


# Label by Strain.Type (WT or KO)
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "Strain.Type")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by Strain.Type") +
  theme_light()


# Label by Analysis.Plate.96_Replicate
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "Analysis.Plate.96_Replicate")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by Analysis.Plate.96_Analysis.Plate.384 - complete") +
  theme_light()


# Label by Wrong_batch
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "Wrong_batch")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by Wrong_batch - full data") +
  theme_light()
```
Found what is causing the outliers, it's batch 3_2 :D


Plot PCA - by biological variables - WTs are still included, so there are NAs in the plots due to that
```{r}
# Label by chromosome


# Label by anticodon
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "Anticodon")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by anticodon") +
  theme_light()


# Label by amino acid
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "Best_isotype_model")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)
#PC <- PC %>% filter(Label != "WT")

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by amino acid") +
  theme_light()


# Label by intron in the tRNA gene
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "Intron")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)
#PC <- PC %>% filter(Label != "WT")

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - intron in the tRNA gene") +
  theme_light()


# Label by whether they are iMet or not
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "iMet")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)
#PC <- PC %>% filter(Label != "WT")

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by whether they are iMet or not") +
  theme_light()


# Label by gene seq length
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "length_DNA_seq")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)
#PC <- PC %>% filter(Label != "WT")

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by gene seq length") +
  theme_light()


# Label by mature seq length
PCA_tags <- get_PCA_tags(PCA_data, sample_layout, "length_mature_seq")
PC <- as.data.frame(results_PCA$x)
PC$Label<- as.factor(PCA_tags)
#PC <- PC %>% filter(Label != "WT")

ggplot(data = PC, aes(x = PC1, y = PC2, col = Label)) +
  geom_point() +
  xlab(glue("PC1 - {var_exp_PC1}%")) +
  ylab(glue("PC2 - {var_exp_PC2}%")) +
  labs(title = "PCA plot - by mature seq length") +
  theme_light()
```
None of these seem to separate anything in the PCA


Remove unnecessary variables created for PCA
```{r}
rm(PCA_tags, var_exp_PC1, var_exp_PC2, var_explained, get_PCA_tags, PC, PCA_data, results_PCA)
```




# 4. Check stats file to see if that explains the differences in batch 3_2
## 4.1. Load and pre-process stats_file
```{r}
# Add column with subbatches to stats file
stats_file <- left_join(stats_file, sample_layout[,c("File.Name", "Analysis.Plate.96", "Replicate", "Analysis.Plate.96_Replicate", "Strain.Type", 
                                                     "Wrong_batch", "date", "Injection_Order", "Date_Injection_Order")], by = "File.Name")
```


## 4.2. Plots to try to explain what is wrong with the subbatch
```{r}
temp <- stats_file %>%
  mutate(QCs = case_when(Analysis.Plate.96 == "QC" ~"QC",
                         TRUE ~ "Not"))

# Plot precursor number vs. protein number coloring by sub-batch
output_file <- paste(base_dir, "tRNA_KOs/Output/Plots/QC/stats_file/precs_vs_proteins_identified_by_subbatch.png", sep="")
ggplot(data = temp, aes(x = Precursors.Identified, y = Proteins.Identified, col = as.factor(Replicate))) +
  geom_point() +
  theme_light()
ggsave(output_file)

# Plot precursor number vs. protein number, color by whether they are in the wrong batch or not
output_file <- paste(base_dir, "tRNA_KOs/Output/Plots/QC/stats_file/precs_vs_proteins_identified_by_wrong_batch.png", sep="")
ggplot(data = stats_file, aes(x = Precursors.Identified, y = Proteins.Identified, col = Wrong_batch)) +
  geom_point() +
  theme_light()
ggsave(output_file)

# Plot precursor number vs. date_injection_order, color by sub-batch
output_file <- paste(base_dir, "tRNA_KOs/Output/Plots/QC/stats_file/precs_identified_vs_injection_by_subbatch.png", sep="")
ggplot(data = stats_file, aes(x = Date_Injection_Order, y = Precursors.Identified, col = Analysis.Plate.96_Replicate)) +
  geom_point() +
  theme_light() +
  theme(axis.text.x=element_blank(), 
        axis.ticks.x=element_blank(),
        # Remove the vertical grid lines
        panel.grid.major.x = element_blank())
ggsave(output_file)

# Plot protein number vs. date_injection_order, color by sub-batch
output_file <- paste(base_dir, "tRNA_KOs/Output/Plots/QC/stats_file/precs_vs_proteins_identified_by_subbatch.png", sep="")
ggplot(data = stats_file, aes(x = Date_Injection_Order, y = Proteins.Identified, col = Analysis.Plate.96_Replicate)) +
  geom_point() +
  theme_light() +
  theme(axis.text.x=element_blank(), 
        axis.ticks.x=element_blank(),
        # Remove the vertical grid lines
        panel.grid.major.x = element_blank())
ggsave(output_file)

# Plot precursor number vs. date_injection_order, color by date
output_file <- paste(base_dir, "tRNA_KOs/Output/Plots/QC/stats_file/precs_identified_vs_injection_by_date.png", sep="")
ggplot(data = stats_file, aes(x = Date_Injection_Order, y = Precursors.Identified, col = as.factor(date))) +
  geom_point() +
  theme_light() +
  theme(axis.text.x=element_blank(), 
        axis.ticks.x=element_blank(),
        # Remove the vertical grid lines
        panel.grid.major.x = element_blank())
ggsave(output_file)

# Plot precursor number vs. MS1.Signal coloring by sub-batch
ggplot(data = stats_file, aes(x = MS1.Signal, y = Precursors.Identified, col = Analysis.Plate.96_Replicate)) +
  geom_point() +
  theme_light()
```


## 4.3. Plots to try to explain what's up with the 2 separate curves in Precursors.Identified vs. Proteins.Identified
```{r}
# Precs vs. proteins
ggplot(data = stats_file, aes(x = Precursors.Identified, y = Proteins.Identified)) +
  geom_point() +
  theme_light()

# Precs vs. proteins, color by batch/384-well plate
ggplot(data = stats_file, aes(x = Precursors.Identified, y = Proteins.Identified, col = as.factor(Replicate))) +
  geom_point() +
  theme_light()

# Precs vs. proteins, color by 96-well plate
ggplot(data = stats_file, aes(x = Precursors.Identified, y = Proteins.Identified, col = Analysis.Plate.96)) +
  geom_point() +
  theme_light()

# Precs vs. proteins, color by 96-well plate_Replicate
ggplot(data = stats_file, aes(x = Precursors.Identified, y = Proteins.Identified, col = Analysis.Plate.96_Replicate)) +
  geom_point() +
  theme_light()

# Precs vs. proteins, color by date
ggplot(data = stats_file, aes(x = Precursors.Identified, y = Proteins.Identified, col = as.factor(date))) +
  geom_point() +
  theme_light()

# Precs vs. proteins, color by KO or WT
ggplot(data = stats_file, aes(x = Precursors.Identified, y = Proteins.Identified, col = Strain.Type)) +
  geom_point() +
  theme_light()

# Precs vs. proteins, color by Date_Injection_Order - doesn't really work, needs to be a factor and it has way too many levels
#ggplot(data = stats_file, aes(x = Precursors.Identified, y = Proteins.Identified, col = Date_Injection_Order)) +
#  geom_point() +
#  theme_light()

# Precs vs. proteins, color by Injection_Order
ggplot(data = stats_file, aes(x = Precursors.Identified, y = Proteins.Identified, col = Injection_Order)) +
  geom_point() +
  theme_light()

# Precs vs. proteins, color by Date+Injection_Order
ggplot(data = stats_file, aes(x = Precursors.Identified, y = Proteins.Identified, col = date+Injection_Order)) +
  geom_point() +
  theme_light()
```
Nothing figured out from this yet :(


## 4.4. Look at this again, but based on the corrected data
Prepare data
```{r}
# Get a dataframe with the number of precursors and proteins and so on, but from the processed data
## Precursor level
report <- report %>%
  dplyr::select(-c(Precursor.Id, Stripped.Sequence, Protein.Group, Genes))

Total.Precursor.Sum <- c()
Precursor.Count <- c()
Sample.ID.unique <- str_replace_all(colnames(report), "-", "_")

for (i in 1:ncol(report)) {
  Total.Precursor.Sum <- c(Total.Precursor.Sum, sum(report[,i], na.rm = T))
  Precursor.Count <- c(Precursor.Count, sum(!is.na(report[,i])))
}

stats_file_processed <- data.frame(Total.Precursor.Sum, Precursor.Count, Sample.ID.unique) %>%
  left_join(sample_layout, by = "Sample.ID.unique")


## Protein level
Total.Protein.Sum <- c()
Protein.Count <- c()
raw_proteomics_colnames <- c()
for (i in 1:ncol(proteomics_processed)) {
  Total.Protein.Sum <- c(Total.Protein.Sum, sum(proteomics_processed[,i], na.rm = T))
  Protein.Count <- c(Protein.Count, sum(!is.na(proteomics_processed[,i])))
  raw_proteomics_colnames <- c(raw_proteomics_colnames, colnames(proteomics_processed)[i])
}

temp <- data.frame(Total.Protein.Sum, Protein.Count, raw_proteomics_colnames)

stats_file_processed <- left_join(temp, stats_file_processed, by = "raw_proteomics_colnames")
```

Plots
```{r}
ggplot(data = stats_file_processed, aes(x = Total.Precursor.Sum, y = Total.Protein.Sum)) +
  geom_point() +
  theme_light()

ggplot(data = stats_file_processed, aes(x = Precursor.Count, y = Protein.Count)) +
  geom_point() +
  theme_light() +
  xlab("Precursors.Identified") +
  ylab("Proteins.Identified")
```


## 4.5. Obtain plots Ludwig recommended, log10(intensity) vs. acquisition time
```{r}
# This code is from 2 chunks above
report <- report %>%
  t() %>%
  as.data.frame()
report$Sample.ID <- as.character(str_replace_all(rownames(report), "-", "_"))
temp <- sample_layout %>%
  dplyr::select(Sample.ID, final_proteomics_colnames, Date_Injection_Order)

report <- left_join(report, temp, by = "Sample.ID")

report <- pivot_longer(data = report, 
                       cols = !c(Sample.ID, final_proteomics_colnames, Date_Injection_Order),
                       names_to = "Whatever",
                       values_to = "Precursor.Quantity")


# Get the plot
temp <- report[1:4701996,]
ggplot(data = temp, aes(x = Date_Injection_Order, y = log10(Precursor.Quantity))) +
  geom_boxplot(outliers = F) +
  theme_light() +
  theme(panel.grid.major.x = element_blank(),
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank()) +
  labs(title = "Total precursor intensity along the injection order, for processed data",
       subtitle = "Only the first 1/5 of the data")
```


These are the proteins that were drift corrected, in case I want to check them
```{r}
drift <- as.data.frame(fread("S:/AG/AG-CF-HTMS/AG-Ralser-Share/30-0092_AndreaLehmann-AlternativeAAUsage-tRNA/05_DataAnalysis/11_Preprocessing_Boris/AlternativeAAUsage-tRNA/driftcorrection_r2_df.tsv"))
```













############################## FROM HERE IT'S OLD CODE - CHECK IT AT SOME POINT ##############################


## 2.2. Clustering analysis
```{r}
# Elbow method - try to figure out what number of clusters is appropriate 
clust_df = log2(trna_ko)
clust_df = as.data.frame(t(clust_df))         
clust_df = scale(clust_df)

fviz_nbclust(clust_df, kmeans, method = "wss", k.max = 15)


# Create a function that can be used to perform clustering with different k values 
perform_clustering = function(tmp, dist_method, clust_method, num_of_clusters, want_dendrogram = T, want_clust_plot = T) {
  # Get distance matrix based on Euclidean distance
  dist.euc = dist(tmp, method = dist_method)
  
  # Perform clustering
  tree.complete = hclust(dist.euc, method = clust_method)
  
  # Plot dendrogram
  if (want_dendrogram) {
    plot(tree.complete, cex = 0.3, 
    main = glue("Dendrogram for clustering with {clust_method} link"),
    xlab = glue("{dist_method} distance"))
    rect.hclust(tree.complete , k = num_of_clusters)
  }
  
  if (want_dendrogram) {
    dend_obj = as.dendrogram(tree.complete)
    col_dend <- color_branches(dend_obj, k = num_of_clusters)
    labels_cex(col_dend)
    plot(col_dend)
  }
  
  # Visualize clusters
  my_clusters = cutree(tree.complete, k = num_of_clusters)
  if (want_clust_plot) {print(fviz_cluster(list(data = tmp, cluster = my_clusters), main = glue("Clustering by {clust_method} link")))}
  
  return(my_clusters)
}


# Use this function to do clustering: k = 2
my_clusters = perform_clustering(clust_df, "euclidean", "complete", 2)
```
--> Maybe I should do this with the data after subtracting the WT median from them or something like that??? That might allow to cluster nicely the major and minor tRNA copy KOs.




## 2.3. Heatmaps
Heatmap on the original log2 proteomics data
```{r}
# Get the heatmap
mat = as.matrix(log2(trna_ko))

col_fun = colorRamp2(c(5, 12.5, 20), c("blue", "white", "red"))     # Not using this now but this is how you do it

Heatmap(mat,
        name = "Heatmap",
        col = col_fun,
        na_col = "black",
        row_title = "Proteins",
        column_title = "Samples",
        #cluster_rows = FALSE,                  # Turn row clustering off, I assume we already know the relationships between the genes
        column_dend_height = unit(2, "cm"), 
        show_row_names = FALSE,                
        column_names_gp = gpar(fontsize = 5))
```

Heatmap with z-score based on median per row
```{r}
mat_zscore = t(apply(log2(trna_ko), 1, function(x) (x-median(x, na.rm = T))/mad(x, na.rm = T)))
col_fun = colorRamp2(c(-4, -2, 0, 2, 4), c("blue", "blue", "white", "red", "red"))

Heatmap(mat_zscore,
        name = "Heatmap",
        #col = col_fun,
        na_col = "green",
        row_title = "Proteins",
        column_title = "Samples",
        #column_labels = grepl("QC", colnames(mat_zscore)),
        #cluster_rows = FALSE,                  # Turn row clustering off, I assume we already know the relationships between the genes
        column_dend_height = unit(2, "cm"), 
        show_row_names = FALSE,                
        column_names_gp = gpar(fontsize = 5))
```

Heatmap with per-gene log2 FC across strains (compared to median of gene across all strains)
```{r}
temp = t(apply(log2(trna_ko), 1, function(x) x/median(x, na.rm = T)))
mat_log2FC = log2(temp)
col_fun = colorRamp2(c(-2, -1, 0, 1, 2), c("blue", "blue", "white", "red", "red"))

Heatmap(mat_log2FC,
        name = "Heatmap",
        col = col_fun,
        na_col = "green",
        row_title = "Proteins",
        column_title = "Samples",
        #cluster_rows = FALSE,                  # Turn row clustering off, I assume we already know the relationships between the genes
        column_dend_height = unit(2, "cm"), 
        show_row_names = FALSE,                # When I use this, rows are not ordered by abundance anymore :(
        column_names_gp = gpar(fontsize = 5))
```
































