---
title: "enrichment_analysis"
author: "Álvaro Gómez Pérez"
date: "2025-02-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Libraries
```{r}
library(clusterProfiler)
library(org.Sc.sgd.db)
library(data.table)
library(dplyr)
library(fgsea)
library(ggplot2)
```



# 1. Load and prepare data
## 1.0. Set parameters
```{r}
# Significance level to be used for all tests and plots in this file
alpha <- 0.01

# Log-fold change limit to be considered "biologically significant"
lfc_threshold <- 1.5

# Directories to be used
working_from = "home"

if (working_from == "home") {
  base_dir = "/home/alvaro/MyStuff/tRNA_KOs/"
} else
if (working_from == "charite") {
  base_dir = "C:/MyStuff/tRNA_KOs/"
}
```

## 1.1. Load data
```{r}
# List of dataframes with the proteins that were DE in each KO strain - at alpha = 0.01
de_proteins_list_001 <- readRDS(paste(base_dir, "Data/Other/enrichment_analysis/de_proteins_list_001.RData", sep=""))

# List of dataframes with the proteins that were DE in each KO strain - at alpha = 0.05
de_proteins_list_005 <- readRDS(paste(base_dir, "Data/Other/enrichment_analysis/de_proteins_list_005.RData", sep=""))

# Full Entrez ID dataframe, with all genes
entrez_db <- as.data.frame(fread(paste(base_dir, "Data/Other/enrichment_analysis/entrez_reference.txt", sep="")))

# Load the dataset with that information
master_dataset <- as.data.frame(fread(paste(base_dir, "Data/Other/GtRNAdb/master_tRNA_dataset.csv", sep="")))

# Phenotypic data from Bloom-Ackermann et al., 2014
phenotypic_data <- as.data.frame(fread(paste(base_dir, "Data/Other/Articles/bloom_ackermann_2014/phenotypic_results.tsv", sep="")))
```



# 2. GO enrichment analysis - for proteins at 0.01
## 2.1. Get NCBI gene IDs for each of the systematic name vectors I have
```{r}
de_proteins_list_entrez_ID_001 <- list()
for (i in 1:length(de_proteins_list_001)) {
  temp <- data.frame("Protein_stable_ID" = de_proteins_list_001[[i]])
  temp <- left_join(temp, entrez_db, by = "Protein_stable_ID")
  temp <- temp$`NCBI_gene_(formerly_Entrezgene)_ID`
  de_proteins_list_entrez_ID_001[[i]] <- temp
}
names(de_proteins_list_entrez_ID_001) <- names(de_proteins_list_001)
```

## 2.2. Get the information of how many DE proteins there are for each tRNA KO strain
```{r}
de_prots_per_strain <- c()
for (i in 1:length(de_proteins_list_001)) {
  temp <- de_proteins_list_001[[i]]
  de_prots_per_strain <- c(de_prots_per_strain, length(temp))
}

de_prots_per_strain_001 <- data.frame(Strain.Name = names(de_proteins_list_001),
                                      nDEP = de_prots_per_strain) %>%
  filter(Strain.Name != "WT")
```


## 2.3. Perform enrichment analysis
### Biological process
```{r}
# Come up with background (all proteins detected)
my_universe <- as.character(de_proteins_list_entrez_ID_001[["WT"]])

# Create empty list for GO results, perform GO for each KO strain, and store them there
go_results_list_001_bp <- list()
for (i in 2:length(de_proteins_list_entrez_ID_001)) {
  proteins_detected <- as.character(de_proteins_list_entrez_ID_001[[i]])
  if (length(proteins_detected) > 0) {
    go_results <- enrichGO(gene = proteins_detected, 
                           OrgDb = "org.Sc.sgd.db", 
                           keyType = "ENTREZID", 
                           ont = "BP", 
                           universe = my_universe)
    go_results <- as.data.frame(go_results)
    if (nrow(go_results) > 0) {
      go_results_list_001_bp[[names(de_proteins_list_entrez_ID_001)[i]]] <- go_results
    }
  }
}

# Check how much correlation there is between strains having many DEPs and strains showing some GO enrichment results
has_go_results <- c()
number_of_go_terms <- c()
for (i in 1:nrow(de_prots_per_strain_001)) {
  strain <- de_prots_per_strain_001$Strain.Name[i]
  if (strain %in% names(go_results_list_001_bp)) {
    has_go_results <- c(has_go_results, "Yes")
    number_of_go_terms <- c(number_of_go_terms, nrow(go_results_list_001_bp[[strain]]))
  }
  else {
    has_go_results <- c(has_go_results, "No")
    number_of_go_terms <- c(number_of_go_terms, 0)
  }
}
de_prots_per_strain_001$GO_results_bp <- has_go_results
de_prots_per_strain_001$num_of_go_terms_bp <- number_of_go_terms
```

### Molecular function
```{r}
# Come up with background (all proteins detected)
my_universe <- as.character(de_proteins_list_entrez_ID_001[["WT"]])

# Create empty list for GO results, perform GO for each KO strain, and store them there
go_results_list_001_mf <- list()
for (i in 2:length(de_proteins_list_entrez_ID_001)) {
  proteins_detected <- as.character(de_proteins_list_entrez_ID_001[[i]])
  if (length(proteins_detected) > 0) {
    go_results <- enrichGO(gene = proteins_detected, 
                           OrgDb = "org.Sc.sgd.db", 
                           keyType = "ENTREZID", 
                           ont = "MF", 
                           universe = my_universe)
    go_results <- as.data.frame(go_results)
    if (nrow(go_results) > 0) {
      go_results_list_001_mf[[names(de_proteins_list_entrez_ID_001)[i]]] <- go_results
    }
  }
}

# Check how much correlation there is between strains having many DEPs and strains showing some GO enrichment results
has_go_results <- c()
number_of_go_terms <- c()
for (i in 1:nrow(de_prots_per_strain_001)) {
  strain <- de_prots_per_strain_001$Strain.Name[i]
  if (strain %in% names(go_results_list_001_mf)) {
    has_go_results <- c(has_go_results, "Yes")
    number_of_go_terms <- c(number_of_go_terms, nrow(go_results_list_001_mf[[strain]]))
  }
  else {
    has_go_results <- c(has_go_results, "No")
    number_of_go_terms <- c(number_of_go_terms, 0)
  }
}
de_prots_per_strain_001$GO_results_mf <- has_go_results
de_prots_per_strain_001$num_of_go_terms_mf <- number_of_go_terms
```

### Cellular compartment
```{r}
# Come up with background (all proteins detected)
my_universe <- as.character(de_proteins_list_entrez_ID_001[["WT"]])

# Create empty list for GO results, perform GO for each KO strain, and store them there
go_results_list_001_cc <- list()
for (i in 2:length(de_proteins_list_entrez_ID_001)) {
  proteins_detected <- as.character(de_proteins_list_entrez_ID_001[[i]])
  if (length(proteins_detected) > 0) {
    go_results <- enrichGO(gene = proteins_detected, 
                           OrgDb = "org.Sc.sgd.db", 
                           keyType = "ENTREZID", 
                           ont = "CC", 
                           universe = my_universe)
    go_results <- as.data.frame(go_results)
    if (nrow(go_results) > 0) {
      go_results_list_001_cc[[names(de_proteins_list_entrez_ID_001)[i]]] <- go_results
    }
  }
}

# Check how much correlation there is between strains having many DEPs and strains showing some GO enrichment results
has_go_results <- c()
number_of_go_terms <- c()
for (i in 1:nrow(de_prots_per_strain_001)) {
  strain <- de_prots_per_strain_001$Strain.Name[i]
  if (strain %in% names(go_results_list_001_cc)) {
    has_go_results <- c(has_go_results, "Yes")
    number_of_go_terms <- c(number_of_go_terms, nrow(go_results_list_001_cc[[strain]]))
  }
  else {
    has_go_results <- c(has_go_results, "No")
    number_of_go_terms <- c(number_of_go_terms, 0)
  }
}
de_prots_per_strain_001$GO_results_cc <- has_go_results
de_prots_per_strain_001$num_of_go_terms_cc <- number_of_go_terms
```


## 2.4. Add other informational columns to this dataset that might explain the success of the GO analysis for some strains
```{r}
de_prots_per_strain_001 <- left_join(de_prots_per_strain_001, master_dataset, by = "Strain.Name")
```


## 2.5. Check the enriched pathways for the KO strains that were checked for growth under DTT in the original article
Biological process
```{r}
## Grew ~ as WT under DTT
View(go_results_list_001_bp[["tR(UCU)M2"]])
View(go_results_list_001_bp[["tH(GUG)G1"]])

## Grew better than WT under DTT
View(go_results_list_001_bp[["tR(CCU)J"]])
View(go_results_list_001_bp[["tL(GAG)G"]])
View(go_results_list_001_bp[["tM(CAU)C"]])
```

Cellular compartment
```{r}
## Grew ~ as WT under DTT
View(go_results_list_001_cc[["tR(UCU)M2"]])
View(go_results_list_001_cc[["tH(GUG)G1"]])

## Grew better than WT under DTT
View(go_results_list_001_cc[["tR(CCU)J"]])
View(go_results_list_001_cc[["tL(GAG)G"]])
View(go_results_list_001_cc[["tM(CAU)C"]])
```

Molecular function
```{r}
## Grew ~ as WT under DTT
View(go_results_list_001_mf[["tR(UCU)M2"]])
View(go_results_list_001_mf[["tH(GUG)G1"]])

## Grew better than WT under DTT
View(go_results_list_001_mf[["tR(CCU)J"]])
View(go_results_list_001_mf[["tL(GAG)G"]])
View(go_results_list_001_mf[["tM(CAU)C"]])
```




# 3. GO enrichment analysis - for proteins at 0.05
## 3.1. Get NCBI gene IDs for each of the systematic name vectors I have
```{r}
de_proteins_list_entrez_ID_005 <- list()
for (i in 1:length(de_proteins_list_005)) {
  temp <- data.frame("Protein_stable_ID" = de_proteins_list_005[[i]])
  temp <- left_join(temp, entrez_db, by = "Protein_stable_ID")
  temp <- temp$`NCBI_gene_(formerly_Entrezgene)_ID`
  de_proteins_list_entrez_ID_005[[i]] <- temp
}
names(de_proteins_list_entrez_ID_005) <- names(de_proteins_list_005)
```

## 3.2. Get the information of how many DE proteins there are for each tRNA KO strain
```{r}
de_prots_per_strain <- c()
for (i in 1:length(de_proteins_list_005)) {
  temp <- de_proteins_list_005[[i]]
  de_prots_per_strain <- c(de_prots_per_strain, length(temp))
}

de_prots_per_strain_005 <- data.frame(Strain.Name = names(de_proteins_list_005),
                                      nDEP = de_prots_per_strain) %>%
  filter(Strain.Name != "WT")
```

## 3.3. Perform enrichment analysis
```{r}
# Come up with background (all proteins detected)
my_universe <- as.character(de_proteins_list_entrez_ID_005[["WT"]])

# Create empty list for GO results, perform GO for each KO strain, and store them there
go_results_list_005 <- list()
for (i in 2:length(de_proteins_list_entrez_ID_005)) {
  proteins_detected <- as.character(de_proteins_list_entrez_ID_005[[i]])
  if (length(proteins_detected) > 0) {
    go_results <- enrichGO(gene = proteins_detected, 
                           OrgDb = "org.Sc.sgd.db", 
                           keyType = "ENTREZID", 
                           ont = "BP", 
                           universe = my_universe)
    go_results <- as.data.frame(go_results)
    if (nrow(go_results) > 0) {
      go_results_list_005[[names(de_proteins_list_entrez_ID_005)[i]]] <- go_results
    }
  }
}

```




# 4. Compare EA results to growth in DTT (equal or worse than WT vs. better than WT) 
## 4.1. Evaluate growth in DTT and classify strains to better than WT or equal/worse
```{r}
# Plot GR and GY in DTT
ggplot(data = phenotypic_data, aes(x = DTT)) +
  geom_histogram(binwidth = 1) +
  theme_light() +
  geom_vline(xintercept = mean(phenotypic_data$DTT, na.rm = T)+sd(phenotypic_data$DTT, na.rm = T), col = "red") +
  geom_vline(xintercept = mean(phenotypic_data$DTT, na.rm = T)-sd(phenotypic_data$DTT, na.rm = T), col = "red")

ggplot(data = phenotypic_data, aes(x = DTT_GY)) +
  geom_histogram(binwidth = 1) +
  theme_light() +
  geom_vline(xintercept = mean(phenotypic_data$DTT_GY, na.rm = T)+sd(phenotypic_data$DTT_GY, na.rm = T), col = "red") +
  geom_vline(xintercept = mean(phenotypic_data$DTT_GY, na.rm = T)-sd(phenotypic_data$DTT_GY, na.rm = T), col = "red")

# For now I am just going to take 0 as the threshold, since these data are in comparison to WT, idk if they published the raw data, I don't think so
# In any case maybe some more restrictive thresholds (like the SDs I plotted above) would be better, but for now this will do

# Need to look into the original article to figure out which one to use - going with both for now, so I don't take all strains here, only those that agree on their GR and GY (not sure if that makes sense, going by vibes for now)
strains_better_in_DTT <- phenotypic_data$gene_name[phenotypic_data$DTT > 0 & phenotypic_data$DTT_GY > 0]
strains_worse_or_equal_in_DTT <- phenotypic_data$gene_name[phenotypic_data$DTT <= 0 & phenotypic_data$DTT_GY <= 0]
```

## 4.2. Compare EA results to these strain classification
Biological process
```{r}
go_list_better_dtt <- list()
go_list_worse_dtt <- list()

for (i in 1:length(go_results_list_001_bp)) {
  strain_name <- names(go_results_list_001_bp)[i]
  if (strain_name %in% strains_better_in_DTT) {
    go_list_better_dtt[[strain_name]] <- go_results_list_001_bp[[i]]
  }
  else if (strain_name %in% strains_worse_or_equal_in_DTT) {
    go_list_worse_dtt[[strain_name]] <- go_results_list_001_bp[[i]]
  }
}
```






# 5. The one Alexis did
```{r}
# Load data
da_full <- readRDS(paste(base_dir, "Data/Other/enrichment_analysis/da.RData", sep=""))

# Alexis' code: EA for GO, KEGG and TF all at once, and summarize which gene sets are enriched in the most tRNA KOs
ea <- list()
for (i in unique(da_full$Strain.Name)) {
  ea[[i]] <- pull(filter(da_full, Strain.Name %in% i & abs(logFC) >= log2(lfc_threshold) & adj.P.Val <= alpha), protein)
}
rm(i)

ea <- gost(ea, organism = 'scerevisiae', correction_method = 'fdr', domain_scope = 'custom', custom_bg = unique(da_full$protein), sources = c('GO', 'KEGG', 'TF'))

ea_terms <- ea$result %>%
  group_by(query) %>%
  summarise(n_terms = n()) %>%
  mutate(AA = str_sub(query, 2, 2),
         chromosome = str_extract(str_remove(query, '[:digit:]$'), '[:alpha:]$'),
         anticodon = str_extract(query, '(?<=\\.)[:alpha:]{3}'))

ea_terms_ko <- ea$result %>%
  group_by(term_name) %>%
  summarise(n_kos = n(), source = unique(source)) %>%
  ungroup() %>%
  filter(n_kos > 25)
```






# 6. GSEA - try to imitate the results from the original article
## 6.1. Prepare ranks
```{r}
# Load list of all dataframes (one per strain) wiht the DE results
da <- readRDS(paste(base_dir, "Data/Other/enrichment_analysis/da.RData", sep=""))
da <- da[names(da) != "WT"]

# Create empty list to put the named ranked gene lists in
ranked_genes <- list()

# Iterate over these dataframes
for (i in 1:length(da)) {
  temp <- da[[i]]
  
  # Since we are using all genes, not only DE ones, we need to rank them
  rankings <- sign(temp$logFC)*(-log10(temp$adj.P.Val))
  names(rankings) <- rownames(temp)
  rankings <- sort(rankings, decreasing = TRUE)
  
  # Add them to the list where we are saving them
  ranked_genes[[names(da)[i]]] <- rankings
  
  # Plot
  plot(rankings)
  
  # Check max and min - to see if there is a need to correct for -Inf and +Inf
  max(rankings)
  min(rankings)
}

```

## 6.2. Perform GSEA
```{r}

```
















## 2.5. THIS IS WHAT I HAD WRITTEN ABOVE BEFORE (WITHIN GO ENRICHMENT ANALYSIS) BEFORE REALIZING I NEEDED TO DO GSEA SEPARATELY, JUST IN CASE I CAN REUSE THE TEXT, EVEN IF NOT THE CODE

### For SC strains: pathways related to proteotoxic stress were significantly induced w.r.t. to the MC group!! That is not what I am checking here, keep that in mind. These include:
  - Proteasome
  - Protein processing in endoplasmic reticulum
```{r}
SC_strains <- phenotypic_data$gene_name[phenotypic_data$MC_or_SC == "SC"]

for (i in 1:length(go_results_list_001)) {
  if (names(go_results_list_001)[i] %in% SC_strains) {
    print(go_results_list_001[[i]][["Description"]])
  }
}
```

### For MC strains: pathways related to translation were significantly induced w.r.t. to the SC group!! That is not what I am checking here, keep that in mind. These include:
  - Ribosome biogenesis
  - Ribosome
```{r}
MC_strains <- phenotypic_data$gene_name[phenotypic_data$MC_or_SC == "MC"]

for (i in 1:length(go_results_list_001)) {
  if (names(go_results_list_001)[i] %in% MC_strains) {
    print(go_results_list_001[[i]][["Description"]])
  }
}
```


































