---
title: "try_limpa_reanalysis"
author: "Álvaro Gómez Pérez"
date: "2025-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




############################################################################################################
########################################## CONTENTS OF THIS FILE ###########################################
############################################################################################################
I only run the first part of limpa here, the protein quantification from the precursor-level data, save that, and load it in the main script, where I perform the DE analysis





Packages
```{r}
library(tidyr)
library(dplyr)
library(data.table)
library(ggplot2)
library(limpa)
library(stringr)
library(helperfunctions)
```

Set up
```{r}
# Significance level to be used for all tests and plots in this file
alpha <- 0.05

# Significance level as a plain string to use when loading or writing files
alpha_plain <- str_replace(as.character(alpha), "\\.", "")

# Minimum threshold for the average log2-expression across all samples in order to keep a protein in the data before DE analysis
mean_log2_across_all_samples_threshold <- 2

# Minimum threshold for the variance of the log2-expression across all samples in order to keep a protein in the data before DE analysis
var_across_log2_all_samples_threshold <- 0.6

# Separate significance level, the one used for the enrichment analysis
alpha_enrichment <- 0.05

# Significance level above as a plain string to use when loading or writing files
alpha_enrichment_plain <- str_replace(as.character(alpha_enrichment), "\\.", "")

# Log-fold change limit to be considered "biologically significant"
lfc_threshold <- 1

# Number of nDEP from which a tRNA is considered "major" (this threshold value should be included) - THIS I AM NOT USING ANYMORE, AM I?
threshold_major <- 2
```

Load data
```{r}
# Boris pre-processed data - normalized, batch and drift corrected
raw_data <- as.data.frame(fread("/home/alvaro/MyStuff/Projects/tRNA_KOs/Data/proteomics_data/AlternativeAAUsage-tRNA_peptidecentric_PrecursorQuantity_reprep_driftcorrected_batchcorrected_onWTs.tsv"))
rownames(raw_data) <- raw_data$Precursor.Id

# Sample layout - from the S drive
metadata <- as.data.frame(fread("/home/alvaro/MyStuff/Projects/tRNA_KOs/Data/basic/metadata_alvaro.tsv"))

# Separate metadata
metadata_columns <- raw_data %>%
  dplyr::select(Precursor.Id, Stripped.Sequence, Protein.Group, Genes)
rownames(metadata_columns) <- NULL
raw_data <- raw_data %>%
  dplyr::select(-c(Precursor.Id, Stripped.Sequence, Protein.Group, Genes))

# Fix colnames (sample names)
colnames(raw_data) <- sapply(colnames(raw_data), function(x) str_replace_all(x, "-", "_"))
temp_match <- metadata %>%
  dplyr::select(Sample.ID.unique, final_proteomics_colnames)
temp_colnames <- data.frame(Sample.ID.unique = colnames(raw_data)) %>%
  left_join(temp_match, by = "Sample.ID.unique")
colnames(raw_data) <- temp_colnames$final_proteomics_colnames

# UniProt dataset
uniprot_db <- as.data.frame(fread("/home/alvaro/MyStuff/Projects/tRNA_KOs/Data/databases/uniprotkb_proteome_UP000002311_2025_05_24.tsv"))

# Remove unnecessary variables
rm(temp_match, temp_colnames)
```


Run limpa protein summarization
```{r}
# Protein labels
protein.id <- metadata_columns$Protein.Group

# Run protein summarization
summarization_results <- protein_summarize(raw_data = raw_data,
                                           protein_ids = protein.id,
                                           method = "limpa")
y.protein <- summarization_results$y.protein
dpcfit <- summarization_results$dpcfit
```


Change protein names - from UniProt IDs to systematic yeast naming
```{r}
# Change the protein names from the UniProt ones (produced by DIA-NN) to systematic Sc ones
## For the rownames of y.protein$E
temp_uniprot <- uniprot_db %>%
  dplyr::select(Entry, `Gene Names (ordered locus)`) %>%
  dplyr::rename(Gene.secondaryIdentifier = `Gene Names (ordered locus)`)
temp_rownames <- data.frame(rownames(y.protein$E)) 
colnames(temp_rownames) <- "Entry"
temp_rownames <- temp_rownames %>%
  left_join(temp_uniprot, by = "Entry")

## For the protein names (they are the same and should be in the same order, but I do it separately just in case), stored in y.protein$genes$Protein
temp_protein_names <- data.frame(y.protein$genes$Protein)
colnames(temp_protein_names) <- "Entry"
temp_protein_names <- temp_protein_names %>%
  left_join(temp_uniprot, by = "Entry")


# Some systematic names contain 2 of them - I have to deal with that - I'M TAKING THE FIRST ONE, HOPE THAT'S NOT TOO BAD
remove_second_systematic_name <- function(name) {
  if (grepl(";", name)) {
    new_name <- substr(name, 0, str_locate(name, ";")[1]-1)
  }
  else {
    new_name <- name
  }
  return(new_name)
}


# Remove the second systematic names wherever it's present and set the new names to the limpa output object to be saved
## For y.protein$E rownames
temp_rownames$Gene.secondaryIdentifier <- sapply(temp_rownames$Gene.secondaryIdentifier, remove_second_systematic_name)
rownames(y.protein$E) <- temp_rownames$Gene.secondaryIdentifier

## For y.protein$genes$Protein
temp_protein_names$Gene.secondaryIdentifier <- sapply(temp_protein_names$Gene.secondaryIdentifier, remove_second_systematic_name)
y.protein$genes$Protein <- temp_protein_names$Gene.secondaryIdentifier
```

Save limpa results
```{r}
# As .rds
saveRDS(dpcfit, file = "/home/alvaro/MyStuff/Projects/tRNA_KOs/Data/limpa_analysis/dpcfit.rds")
saveRDS(y.protein, file = "/home/alvaro/MyStuff/Projects/tRNA_KOs/Data/limpa_analysis/y.protein.rds")
```

Load limpa results 
```{r}
dpcfit <- readRDS("/home/alvaro/MyStuff/Projects/tRNA_KOs/Data/limpa_analysis/dpcfit.rds")
y.protein <- readRDS("/home/alvaro/MyStuff/Projects/tRNA_KOs/Data/limpa_analysis/y.protein.rds")
```




```{r}

```






















############################## OLD CODE ##############################
Keep until I have tested that the functions above work fine



Prepare limpa input
```{r}
# Protein labels
protein.id <- metadata$Protein.Group

# Design matrix
## Get the names of all strains, and set WT as the reference
my_levels <- c()
for (i in 1:ncol(raw_data)) {
  sample_name <- colnames(raw_data)[i]
  strain_name <- metadata$Strain.Name[metadata$final_proteomics_colnames == sample_name]
  if (length(strain_name) > 1) {
    strain_name <- strain_name[1]
  }
  my_levels <- c(my_levels, strain_name)
}
my_levels <- as.factor(my_levels)
my_levels <- relevel(my_levels, ref = "WT")

## Generate design matrix
mm <- model.matrix(~ 0 + my_levels)
colnames(mm) <- levels(my_levels)
```

Run limpa analysis
```{r}
# Original code
dpcfit <- dpc(raw_data)
y.protein <- dpcQuant(raw_data, protein.id, dpc=dpcfit)
```


Evaluate results with the plots from the limpa package
```{r}
plotDPC(dpcfit)
plotMDSUsingSEs(y.protein)
```





















